---
title: "STAT/CS 187 Homework 3"
author: "Solutions"
date: "Spring 2022"
output:
  word_document:
    keep_md: true
    
---
  
### Set Up Your Project and Load Libraries

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE) 

## Set the default size of figures
knitr::opts_chunk$set(fig.width=8, fig.height=5)  

## Load the libraries we will be using
pacman::p_load(tidyverse)

## Changing the default theme to black white
theme_set(theme_bw())

```

### Question 1: Board Game Geek Reviews by Category

In general, we see that ideal diamonds are least likely to have the worst color (J), and more likely to have better colors than the lesser cuts.  However, the trends are not dramatic.  The first plot makes it clear that there are many more diamonds with better cuts, than those with Fair or Good cuts.

```{r 1_load_boardgames}
# Read in the board game data below:
bgs <- read.csv("board games.csv")

# Next, remove any games with the game_category "Other"
bg1 <- bgs %>% 
  
  filter(game_category != "Other")
  

# Then change the rating_group to a factor with ORDERED levels Terrible, Bad, Average, Good, Great
bg1 <- 
  bg1 %>% 
  
  mutate(rating_group = rating_group %>% 
                        as_factor() %>%   # Changes to a factor with levels ordered in the data
                        as.ordered())     # Makes the column an ordered factor
  
str(bg1)
```

#### Part 1a) Side-by-Side Bar Chart

```{r 1a_bg_sbs}
# Create the side-by-side bar chart for rating by category:
bg1 %>% 
  
  ggplot(mapping = aes(x = game_category,
                       fill = rating_group)) +
  
  geom_bar(position = "dodge") + 

  labs(x = "Board Game Category",
       fill = "Game Rating") + 
  
  theme_test() +
  
  theme(legend.position = "bottom") + 
  
  scale_y_continuous(minor_breaks = NULL,
                     expand = expansion(mult = 0,
                                        add = c(0, 50))) + 
  
  scale_fill_viridis_d(option = "cividis") 
```


#### Part 1b) Conditional Proportions - Rating by Category


``` {r 1b_relprops}
# Create a data set that has the rating proportions for each individual game_category

# Book Method
rating_by_category <- 
  bg1 %>%
  
  group_by(game_category, rating_group) %>%
  
  summarize(rating_cat_n = n()) %>%
  
  mutate(proportion = rating_cat_n / sum(rating_cat_n)) %>% 
  
  dplyr::select(-rating_cat_n)

tibble(rating_by_category)


# Alternative method
rat_by_cat <- 
  
  bg1 %>% 
  
  # Using xtabs to calculate the count for each category and rating combo
  xtabs(formula = ~ rating_group + game_category) %>% 
  
  # prop.table to convert the counts to a conditional proportion
  prop.table(margin = "game_category") %>% 
  
  # Converting the table to a data.frame
  data.frame() %>% 
  
  # Converting rating_group back to an ordered factor
  mutate(rating_group = as.ordered(rating_group)) %>% 
  
  # Changing Freq to proportion
  rename(proportion = Freq) %>% 
  
  # Reordering the columns
  dplyr::select(game_category, rating_group, proportion)

tibble(rat_by_cat)
```

Students only need to use one method above, but should use one of the two. The steps don't need to be identical, but if you see a major deviation, let me know!


#### Part c) Relative Side-by-Side Bar Chart

```{r 1c_rel_sbs}
# Create the Side-by-Side box plot shown in the homework instructions:
ggplot(data = rat_by_cat,
       mapping = aes(x = game_category,
                     y = proportion,
                     fill = rating_group)) + 
  
  geom_col(position = "dodge") + 
  
  theme_test() + 
  
  labs(x = "Board Game Category",
       fill = "Game Rating",
       y = "Percent") +
  
  theme(legend.position = "bottom") +
  
  scale_y_continuous(labels = scales::percent,
                     expand = expansion(mult = 0,
                                        add = c(0, 0.020)))  +
  
  scale_fill_viridis_d(option = "cividis") 
```



###  Question 2: More Board Games - Rating by Year

Question 2 will use the full data set (all 20,000+ board games), not the smaller set from Question 1

#### Part 2a) Average Rating by Year

Create a data set named *games_cat_year* that summarizes the data for each year_published and game_category combination. The data set should have the following columns:

- category = The game_category
- year = The publishing year
- games = the number of games for the category and year combination
- rating_mean = the average rating for all games of the category released for the year
- rating_sdev = the standard deviation for rating_avg for all games of the category released for the year

Only keep the rows from 1990 to the present

Print 

```{r 2a_rating_summ}
# Create the new data.frame 
games_cat_year <- 
  
  bgs %>% 
  
  group_by(game_category, year_published) %>% 
  
  summarize(games = n(),
            rating_mean = mean(rating_avg, na.rm = T),
            rating_sdev = sd(rating_avg, na.rm = T)) %>% 
  
  ungroup() %>% 
  
  rename(category = game_category,
         year = year_published) %>% 
  
  filter(year >= 1990)
  

# Use the code below to print your results
games_cat_year %>%

  group_by(category) %>% 
  
  slice_max(n = 1, 
            order_by = games) %>% 
  
  ungroup() 
  
```

#### Part 2b) Linegraph for Rating by Year per Category



```{r 2b_strategic_linegraph, message=F, warning=F}
gg_strategy_rating <- 
  games_cat_year %>% 
  
  filter(category == "Strategy") %>% 
  
  ggplot(mapping = aes(x = year,
                       y = rating_mean)) + 
  
  geom_line(color = "orchid",
            size = 1) + 
  
  labs(y = "Average Rating",
       x = "Year") 

gg_strategy_rating
```

#### Part 2c) Smoothed Line for Overall Mean by Year

Add a smoothed LOESS line to the previous graph for the average rating for games per year, not separated by category. Set the following aesthetics for the smoothed line:

- size = 0.5
- color = "black"
- linetype = "dashed"

```{r 2c_added_mean_line, message=F, warning=F}
gg_strategy_rating +
  
  geom_smooth(data = games_cat_year %>% 
                   group_by(year) %>% 
                   summarize(overall_avg = sum(rating_mean*games)/sum(games)),
            mapping = aes(y = overall_avg),
            se = F,
            size = 0.5,
            color = "black",
            linetype = "dashed")  +
  
  annotate(geom = "text", label = "Average rating for all games",
           x = 2015, y = 6) + 
  
  annotate(geom = "text", label = "Average rating for Strategy games",
           x = 1995, y = 7.25, color = "orchid")  
  

```


#### Part 2d) Each Game Category

Use small multiples to create the same graph from 2c) for all 9 game categories. 

What can you determine about the opinions of Board Game Geek users about the different category of games?

```{r 2d_allcategories, message=F, warning=F}
games_cat_year %>% 
  
  ggplot(mapping = aes(x = year,
                       y = rating_mean)) + 
  
  geom_line(mapping = aes(color = category),
            size = 1) + 
  
  facet_wrap(facets = ~ category) + 
  
  theme(legend.position = "none") + 
  
  labs(y = "Average Rating",
       x = "Year",
       caption = "Black line represents the average rating for all games") +
  
  geom_smooth(data = games_cat_year %>% 
                   group_by(year) %>% 
                   summarize(overall_avg = sum(rating_mean*games)/sum(games)),
            mapping = aes(y = overall_avg),
            se = F,
            size = 0.5,
            color = "black",
            linetype = "dashed") 



```



### Question 3: Word Data

#### Part a) Change word level order
```{r load_words}
# Read the word data into the R environment
words <- read.csv("words.csv")

# Change word to a factor with the level order as idiot/moron/nimrod/nerd/geek/dork
words <- words %>% 
  
  mutate(word = factor(word,
                        levels = c("idiot", "moron", "nimrod",
                                   "nerd", "geek", "dork")))

# Use the code below to check that the words are in the correct order
levels(words$word)

```



#### 3b) Graph of words over time

```{r 3b_word_time}
# Create a plot of relative usage by year for each word
gg_word_line <- 
  
  words %>% 
  
  ggplot(mapping = aes(x = Year,
                       y = relative_usage,
                       color = word)) + 
  
  geom_line(size = 1) + 
  
  labs(x = "Year",
       y = "Relative Word Usage")

gg_word_line

```


#### 3c) Display Word Without Legend

Instead of using a legend to represent the words, place the word itself at the beginning of each line. The word size is 5


```{r 3c_alternative_legend}
gg_word_line +
  
  theme(legend.position = "none") +
  
  geom_text(data = words %>% 
                   group_by(word) %>% 
                   filter(Year == min(Year)),
            
            mapping = aes(label = word),
            hjust = 1,
            size = 5) + 
  
  # Add the function below to make your plot match what is in the homework description
  scale_x_continuous(limits = c(1840, 2020),
                     breaks = seq(1850, 2000, by = 25),
                     minor_breaks = NULL)

```

