---
title: "STAT/CS 187 Homework 3"
author: "Solutions"
date: "Spring 2022"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
  
### Set Up Your Project and Load Libraries

```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE) 

## Set the default size of figures
knitr::opts_chunk$set(fig.width=8, fig.height=5)  


## Load the libraries we will be using
library(readr)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(lattice)
library(directlabels)
```

### Question 1: Board Game Geek Reviews by Category

See the word doc for the description of the data.

```{r 1_load_boardgames}
# Read in the board game data below:
bgs <-   read_csv("C:/Users/huaye/Desktop/CS 187A/HW/HW_3/board games.csv")
View(bgs)

# Next, remove any games with the game_category "Other" and name it bg1
bg1 <- bgs %>%
  filter(game_category !="Other")
  
View(bg1)


# Then change the rating_group to a factor with ORDERED levels Terrible, Bad, Average, Good, Great

rank<-as.factor(bg1$rating_group)
bg1 <- 
  bg1 %>% mutate(rating_group = factor(rank, ordered = TRUE, levels = c("Terrible","Bad","Average","Good","Great")))

  
# Use the line below to check that rating is a ordered factor  
str(bg1)
```

#### Part 1a) Side-by-Side Bar Chart

```{r 1a_bg_sbs}
# Create the side-by-side bar chart for rating by category:
bg1 %>% group_by(game_category) 

  ggplot(data=bg1, mapping=aes(x=game_category,fill=rating_group))+
  
  geom_bar(position = position_dodge())+

# Add the code below to you gg_object for the graph to match what is in the homework instructions  
  scale_y_continuous(minor_breaks = NULL,
                     expand = expansion(mult = 0,
                                        add = c(0, 50))) + 
  
  scale_fill_viridis_d(option = "cividis") 
```


#### Part 1b) Conditional Proportions - Rating by Category


``` {r 1b_relprops}
# Create a data set that has the rating proportions for each individual game_category

data1<-bgs %>%
  filter(game_category !="Other") %>%
  group_by(game_category,rating_group) %>%
  summarise(n_group=n()) %>%
  mutate(proportion = n_group/sum(n_group))
  
rating_by_category <- data.frame(data1$game_category,data1$rating_group,data1$proportion)  

rank<-as.factor(rating_by_category$data1.rating_group)
rating_by_category <- rating_by_category %>%
  mutate(data1.rating_group = factor(rank, ordered = TRUE, levels = c("Terrible","Bad","Average","Good","Great")))


# Run to check that the data.frame is cone correctly  
tibble(rating_by_category)
```




#### Part c) Relative Side-by-Side Bar Chart

Using the data.frame created in part b), create a plot similar to the barchart in part a), but with the proportion on the y-axis instead of the count

```{r 1c_rel_sbs}
# Create the Side-by-Side box plot shown in the homework instructions:

  ggplot(data=rating_by_category,mapping=aes(x=data1.game_category,y=data1.proportion,fill=data1.rating_group))+
  
  geom_col(position = position_dodge())+


# Similar to 1a), add this code for your graph to match the homework description  
  scale_y_continuous(labels = scales::percent,
                     expand = expansion(mult = 0,
                                        add = c(0, 0.020)))  +
  
  scale_fill_viridis_d(option = "cividis") 
```


#### Part d) Comparing Game Categories

War games tend to have higher ratings (at least by people who use BoardGameGeek.com).

The two graphs in a) and c) look different because the y-axis aredifferent(y-axis for graph a) is count, but y-axis on graph c) is proportion).





###  Question 2: More Board Games - Rating by Year

Question 2 will use the full data set (all 20,000+ board games), not the smaller set from Question 1

#### Part 2a) Average Rating by Year

Create a data set named *games_cat_year* that summarizes the data for each year_published and game_category combination. The data set should have the following columns:

- category = The game_category
- year = The publishing year
- games = the number of games for the category and year combination
- rating_mean = the average rating for all games of the category released for the year
- rating_sdev = the standard deviation for rating_avg for all games of the category released for the year

Only keep the rows from 1990 to the present


```{r 2a_rating_summ}
# Create the new data.frame 
games_cat_year1 <- bgs %>%
  select(game_category,year_published,rating_avg,rating_sd)  %>% 
  rename(category=game_category,year= year_published,rating_mean=rating_avg,rating_sdev= rating_sd) %>%
  filter(year> 1990 | year == 1990) %>%
  group_by(category,year)



games_cat_year2 <- games_cat_year1 %>%
  filter(year> 1990 | year == 1990) %>%
  group_by(category,year) %>%
  summarise(n_game=n()) %>%
  mutate(games=n_game)
  


games_cat_year <- full_join(games_cat_year1,games_cat_year2)%>%
  select(-n_game)#Join dataset and drop column n_game

# Use the code below to print your results. Should match the results in the homework description
games_cat_year %>%

  group_by(category) %>% 
  
  slice_max(n = 1, 
            order_by = games)
  
```

#### Part 2b) Linegraph for Rating by Year per Category

Create a line graph of rating_mean by year only for Strategy games. Make sure that the plot background and labeling looks the same. The line color is "orchid" and has size 1

```{r 2b_strategic_linegraph, message=F, warning=F}
# Create and save the line graph
plot_dat <- games_cat_year %>%
  filter(category %in% c("Strategy")) %>%
  group_by(year,games) 

gg_strategy_rating <- ggplot(data=plot_dat , mapping = aes(x=year,y=rating_mean))+
  geom_line(color="orchid",size=1)+
  xlab("Year")+
  ylab("Average Rating")+
  scale_x_continuous(limits = c(1990, 2020),
                     breaks = seq(1990, 2020, by = 10),
                     minor_breaks = NULL)+
  scale_y_continuous(limits = c(6,7.5),
                     breaks = seq(6, 7.5, by = 0.5),
                     minor_breaks = NULL)
  
  
  distinct(plot_dat,year)

# Print the graph in the knitted document  
gg_strategy_rating
```

#### Part 2c) Smoothed Line for Overall Mean by Year

Add a smoothed LOESS line to the previous graph for the average rating for games per year, not separated by category. Set the following aesthetics for the smoothed line:

- size = 0.5
- color = "black"
- linetype = "dashed"

```{r 2c_added_mean_line, message=F, warning=F}
# Build on the plot created in 2b)
gg_strategy_rating + geom_smooth(method="loess",size = 0.5, color = "black", linetype = "dashed",se=FALSE)
  
  
  

```



#### Part 2d) Each Game Category

Use small multiples to create the same graph from 2c) for all 9 game categories. 

What can you determine about the opinions of Board Game Geek users about the different category of games?

```{r 2d_allcategories, message=F, warning=F}
# Create the small multiple plots seen in the homework description
ggplot(data=games_cat_year,mapping =aes(x=year,y=rating_mean,color=category))+
  geom_line(size=1)+ 
  ylim(5,8)+
  geom_smooth(method="loess",size = 0.5, color = "black", linetype = "dashed",se=FALSE)+
  facet_wrap(~category,nrow=3)+
  xlab("Year")+
  ylab("Average Rating")
  




```

<Your answer here>



### Question 3: Word Data

#### Part a) Change word level order
```{r load_words}
# Read the word data into the R environment
words <- read_csv("C:/Users/huaye/Desktop/CS 187A/HW/HW_3/words.csv")

# Change word to a factor with the level order as idiot/moron/nimrod/nerd/geek/dork
rank<-as.factor(words$word)
words <-  words %>% 
  mutate(word = factor(rank, ordered = TRUE, levels = c("idiot","moron","nimrod","nerd","geek","dork")))


# Use the code below to check that the words are in the correct order
levels(words$word)

```




#### 3b) Graph of words over time

Create the line graph below for relative usage by year. The size of each line is 1

```{r 3b_word_time}
# Create and save a plot of relative usage by year for each word
gg_word_line <- ggplot(data=words,mapping =aes(x=Year,y=relative_usage,,color=word))+
  geom_line(size=1)+ 
  xlab("Year")+
  ylab("Relative Word Lsage")
  
  

# Print the plot in the knitted word document
gg_word_line
```


#### 3c) Display Word Without Legend

Instead of using a legend to represent the words, place the word itself at the beginning of each line. The word size is 5


```{r 3c_alternative_legend}
# Second version of the 3b plot but without the legend.

ggplot(data=words,mapping =aes(x=Year,y=relative_usage,color=word))+
  geom_line(size=1)+ 
  xlab("Year")+
  ylab("Relative Word Lsage")+
  theme(legend.position = "none")+
  directlabels::geom_dl(aes(label = word), method = "smart.grid")+
  

  
  # Add the function below to make your plot match what is in the homework description
  scale_x_continuous(limits = c(1840, 2020),
                     breaks = seq(1850, 2000, by = 25),
                     minor_breaks = NULL)

```

